### 2022-01-05 paper/paper.tex
Allow using pre-generated PDFs from pgfplots
mention number of DOFs
additional performance reference

% Plotting in tikz images
\usepackage{pgfplots}
% Silence warnings
\pgfplotsset{compat=1.16}
% Allow external generation of images to speed up compilation
% \usetikzlibrary{external}
% \tikzexternalize[prefix=../figures/tikz/]

@@

For the problem setup we consider a periodic box of the domain $[-1,1]^3$ with four elements in each
spatial direction. To make the mesh curvilinear, the interior of the box is heavily warped
by a combination of sine and cosine functions~\cite{chan2019efficient}. To integrate up to the
final time $T=1.0$, both codes use the five-stage, fourth-order low-storage explicit
Runge-Kutta method of Carpenter and Kennedy~\cite{CarpenterKennedy1994}.
A stable explicit time step is adaptively computed according to the local maximum wave
speed, the relative grid size, and an adjustable Courant-Friedrichs-Lewy (CFL)
coefficient $\text{CFL}\in(0,1]$~\cite{gassner2011}. For the performance computations presented herein
we fix this coefficient to be $\text{CFL}=0.5$.

@@

\begin{figure}[!ht]
\centering
  \begin{subfigure}{\linewidth}
    \begin{tikzpicture}[
      font=\footnotesize
      ]
      \begin{axis}[
          xlabel={Polynomial degree},
          % ylabel={Time per RHS call and DOF (seconds)},
          ylabel={Time/RHS/DOF {[sec]}},
          width=\textwidth,
          height=0.66\textwidth,
          legend columns=-1,
          % legend to name=place_legend_here,
          legend pos=north west,
          ymin=0.0,
          ymax=2.0e-6,
          grid=major,
        ]
        \addplot[blue, mark=*, solid] % FLUXO, weak form
          table [x index=0, y index=2]{../pid_data/pids_euler_weak.dat};
          \addlegendentry{FLUXO}
        \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
          table [x index=0, y index=1]{../pid_data/pids_euler_weak.dat};
          \addlegendentry{\trixi}
        \node at (axis cs:13,3.5e-7) {Weak form};

        \addplot[blue, mark=*, dashed] % FLUXO, flux differencing
          table [x index=0, y index=2]{../pid_data/pids_euler_ranocha.dat};
        \addplot[red, mark=diamond*, dashed] % StructuredMesh, flux differencing
          table [x index=0, y index=1]{../pid_data/pids_euler_ranocha.dat};
        \node [rotate=15] at (axis cs:9,8.2e-7) {Flux differencing};
        \draw (axis cs:9,9.2e-7) -- (axis cs:8.5,10.2e-7);
        \draw (axis cs:9,7.2e-7) -- (axis cs:8.5,5.2e-7);
      \end{axis}
    \end{tikzpicture}%
    \caption{Absolute run times.}
  \end{subfigure}%
  \\
  \begin{subfigure}{\linewidth}
    \begin{tikzpicture}[
      font=\footnotesize
      ]
      \begin{axis}[
          xlabel={Polynomial degree},
          % ylabel={Time per RHS call and DOF (relative)},
          ylabel={Time relative to FLUXO {[-]}},
          width=\textwidth,
          height=0.66\textwidth,
          legend columns=-1,
          legend pos=south west,
          ymin=0.0,
          ymax=1.1,
          ytick={0, 0.2, 0.4, 0.5, 0.6, 0.8, 1},
          grid=major,
        ]
        \addplot[blue, mark=*, solid, forget plot] % FLUXO, weak form
          table [x index=0, y index=2]{../pid_data/pids_euler_weak_relative.dat};
        \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
          table [x index=0, y index=1]{../pid_data/pids_euler_weak_relative.dat};
          \addlegendentry{Weak form}

@@

\begin{figure}[!ht]
\centering
  \begin{subfigure}{\linewidth}
    \begin{tikzpicture}[
      font=\footnotesize
      ]
      \begin{axis}[
          xlabel={Polynomial degree},
          ylabel={Time/RHS/DOF {[sec]}},
          width=\textwidth,
          height=0.66\textwidth,
          legend columns=-1,
          % legend to name=place_legend_here,
          legend pos=north west,
          ymin=0.0,
          ymax=2.0e-6,
          grid=major,
        ]
        \addplot[blue, mark=*, solid] % FLUXO, weak form
          table [x index=0, y index=2]{../pid_data/pids_mhd_weak.dat};
          \addlegendentry{FLUXO}
        \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
          table [x index=0, y index=1]{../pid_data/pids_mhd_weak.dat};
          \addlegendentry{\trixi}
        \node [rotate=12] at (axis cs:13.5,5.75e-7) {Central flux};
        \draw (axis cs:13,6.0e-7) -- (axis cs:12.4,7.1e-7);
        \draw (axis cs:13.4,6.3e-7) -- (axis cs:13.7,1.075e-6);

        \addplot[blue, mark=*, dashed] % FLUXO, flux differencing
          table [x index=0, y index=2]{../pid_data/pids_mhd_flogor.dat};
        \addplot[red, mark=diamond*, dashed] % StructuredMesh, flux differencing
          table [x index=0, y index=1]{../pid_data/pids_mhd_flogor.dat};
        \node [rotate=20] at (axis cs:7,1.2e-6) {EC flux};
        \draw (axis cs:8.4,1.1e-6) -- (axis cs:7.75,1.15e-6);
        \draw (axis cs:7,1.1e-6) -- (axis cs:7.5,6.5e-7);
      \end{axis}
    \end{tikzpicture}%
    \caption{Absolute run times.}
  \end{subfigure}%
  \\
  \begin{subfigure}{\linewidth}
    \begin{tikzpicture}[
      font=\footnotesize
      ]
      \begin{axis}[
          xlabel={Polynomial degree},
          ylabel={Time relative to FLUXO {[-]}},
          width=\textwidth,
          height=0.66\textwidth,
          legend columns=-1,
          legend pos=south west,
          ymin=0.0,
          ymax=1.1,
          ytick={0, 0.2, 0.5, 0.6, 0.7, 0.8, 0.9, 1},
          grid=major,
        ]
        \addplot[blue, mark=*, solid, forget plot] % FLUXO, weak form
          table [x index=0, y index=2]{../pid_data/pids_mhd_weak_relative.dat};
        \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
          table [x index=0, y index=1]{../pid_data/pids_mhd_weak_relative.dat};
          \addlegendentry{Central flux}

@@

From these results, it is clear that \trixi is more than 2x faster for the compressible Euler equations
and more than 1.5x faster for the ideal MHD equations than the Fortran code FLUXO.
It is also possible to see the additional computational effort for the entropy-conservative flux
differencing DG methods for both physical systems. This is because the numerical fluxes are more
expensive in
terms of computational cost, and require special care to optimize their performance~\cite{hendrik_blog}.
We reiterate that both codes implement the same numerical nodal DG methods on curvilinear meshes used in these tests.
This demonstrates the suitability of Julia for this kind of simulation-focused scientific computing.

While our Julia code is faster than the mature HPC Fortran code FLUXO for this non-trivial
example, we do not claim that Julia is generally faster than Fortran, C, or C++.
Instead, we would like to emphasize that Julia code can be \emph{at least
as fast} as code written in these traditional scientific computing languages.
Further, we expect to be able to achieve similar performance with either language
by spending enough time and effort to optimize the respective codes.
\trixi owes its performance optimizations in part to the code introspection and
profiling tools available in Julia. Similar tools used to optimize the performance in other
languages are often not as easy to use as their Julia counterparts.

>>>>

% Allow external generation of images to speed up compilation
\usepackage{tikz}
\usetikzlibrary{external}
\tikzexternalize[prefix=tikz/]

% Work around the fact that openjournal's whedon does not provide the pgfplots package
% Xref: https://github.com/JuliaCon/proceedings-review/issues/76#issuecomment-913454361
\newif\ifregenerateplots
% \regenerateplotstrue  % Enable this if plots need to be regenerated
\regenerateplotsfalse % Enable this to build PDF with whedon

\ifregenerateplots
% Plotting in tikz images
\usepackage{pgfplots}
% Silence warnings
\pgfplotsset{compat=1.16}
\fi

@@

For the problem setup we consider a periodic box of the domain $[-1,1]^3$ with four elements in each
spatial direction. This results in $64 (\text{\lstinline{polydeg}}+1)^3$ degrees
of freedom (DG nodes) when polynomials of degree \lstinline{polydeg} are used.
To make the mesh curvilinear, the interior of the box is heavily warped
by a combination of sine and cosine functions~\cite{chan2019efficient}. To integrate up to the
final time $T=1.0$, both codes use the five-stage, fourth-order low-storage explicit
Runge-Kutta method of Carpenter and Kennedy~\cite{CarpenterKennedy1994}.
A stable explicit time step is adaptively computed according to the local maximum wave
speed, the relative grid size, and an adjustable Courant-Friedrichs-Lewy (CFL)
coefficient $\text{CFL}\in(0,1]$~\cite{gassner2011}. For the performance computations presented herein
we fix this coefficient to be $\text{CFL}=0.5$.

@@

\begin{figure}[!ht]
\centering
  \begin{subfigure}{\linewidth}
    \tikzsetnextfilename{runtime_euler_absolute}
    \begin{tikzpicture}[
      font=\footnotesize
      ]
      \begin{axis}[
          xlabel={Polynomial degree},
          % ylabel={Time per RHS call and DOF (seconds)},
          ylabel={Time/RHS/DOF {[sec]}},
          width=\textwidth,
          height=0.66\textwidth,
          legend columns=-1,
          % legend to name=place_legend_here,
          legend pos=north west,
          ymin=0.0,
          ymax=2.0e-6,
          grid=major,
        ]
        \addplot[blue, mark=*, solid] % FLUXO, weak form
          table [x index=0, y index=2]{../pid_data/pids_euler_weak.dat};
          \addlegendentry{FLUXO}
        \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
          table [x index=0, y index=1]{../pid_data/pids_euler_weak.dat};
          \addlegendentry{\trixi}
        \node at (axis cs:13,3.5e-7) {Weak form};

        \addplot[blue, mark=*, dashed] % FLUXO, flux differencing
          table [x index=0, y index=2]{../pid_data/pids_euler_ranocha.dat};
        \addplot[red, mark=diamond*, dashed] % StructuredMesh, flux differencing
          table [x index=0, y index=1]{../pid_data/pids_euler_ranocha.dat};
        \node [rotate=15] at (axis cs:9,8.2e-7) {Flux differencing};
        \draw (axis cs:9,9.2e-7) -- (axis cs:8.5,10.2e-7);
        \draw (axis cs:9,7.2e-7) -- (axis cs:8.5,5.2e-7);
      \end{axis}
    \end{tikzpicture}%
    \caption{Absolute run times.}
  \end{subfigure}%
  \\
  \begin{subfigure}{\linewidth}
    \tikzsetnextfilename{runtime_euler_relative}
    \begin{tikzpicture}[
      font=\footnotesize
      ]
      \begin{axis}[
          xlabel={Polynomial degree},
          % ylabel={Time per RHS call and DOF (relative)},
          ylabel={Time relative to FLUXO {[-]}},
          width=\textwidth,
          height=0.66\textwidth,
          legend columns=-1,
          legend pos=south west,
          ymin=0.0,
          ymax=1.1,
          ytick={0, 0.2, 0.4, 0.5, 0.6, 0.8, 1},
          grid=major,
        ]
        \addplot[blue, mark=*, solid, forget plot] % FLUXO, weak form
          table [x index=0, y index=2]{../pid_data/pids_euler_weak_relative.dat};
        \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
          table [x index=0, y index=1]{../pid_data/pids_euler_weak_relative.dat};
          \addlegendentry{Weak form}

@@

\begin{figure}[!ht]
\centering
  \begin{subfigure}{\linewidth}
    \tikzsetnextfilename{runtime_mhd_absolute}
    \begin{tikzpicture}[
      font=\footnotesize
      ]
      \begin{axis}[
          xlabel={Polynomial degree},
          ylabel={Time/RHS/DOF {[sec]}},
          width=\textwidth,
          height=0.66\textwidth,
          legend columns=-1,
          % legend to name=place_legend_here,
          legend pos=north west,
          ymin=0.0,
          ymax=2.0e-6,
          grid=major,
        ]
        \addplot[blue, mark=*, solid] % FLUXO, weak form
          table [x index=0, y index=2]{../pid_data/pids_mhd_weak.dat};
          \addlegendentry{FLUXO}
        \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
          table [x index=0, y index=1]{../pid_data/pids_mhd_weak.dat};
          \addlegendentry{\trixi}
        \node [rotate=12] at (axis cs:13.5,5.75e-7) {Central flux};
        \draw (axis cs:13,6.0e-7) -- (axis cs:12.4,7.1e-7);
        \draw (axis cs:13.4,6.3e-7) -- (axis cs:13.7,1.075e-6);

        \addplot[blue, mark=*, dashed] % FLUXO, flux differencing
          table [x index=0, y index=2]{../pid_data/pids_mhd_flogor.dat};
        \addplot[red, mark=diamond*, dashed] % StructuredMesh, flux differencing
          table [x index=0, y index=1]{../pid_data/pids_mhd_flogor.dat};
        \node [rotate=20] at (axis cs:7,1.2e-6) {EC flux};
        \draw (axis cs:8.4,1.1e-6) -- (axis cs:7.75,1.15e-6);
        \draw (axis cs:7,1.1e-6) -- (axis cs:7.5,6.5e-7);
      \end{axis}
    \end{tikzpicture}%
    \caption{Absolute run times.}
  \end{subfigure}%
  \\
  \begin{subfigure}{\linewidth}
    \tikzsetnextfilename{runtime_mhd_relative}
    \begin{tikzpicture}[
      font=\footnotesize
      ]
      \begin{axis}[
          xlabel={Polynomial degree},
          ylabel={Time relative to FLUXO {[-]}},
          width=\textwidth,
          height=0.66\textwidth,
          legend columns=-1,
          legend pos=south west,
          ymin=0.0,
          ymax=1.1,
          ytick={0, 0.2, 0.5, 0.6, 0.7, 0.8, 0.9, 1},
          grid=major,
        ]
        \addplot[blue, mark=*, solid, forget plot] % FLUXO, weak form
          table [x index=0, y index=2]{../pid_data/pids_mhd_weak_relative.dat};
        \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
          table [x index=0, y index=1]{../pid_data/pids_mhd_weak_relative.dat};
          \addlegendentry{Central flux}

@@

From these results, it is clear that \trixi is more than 2x faster for the compressible Euler equations
and more than 1.5x faster for the ideal MHD equations than the Fortran code FLUXO.
It is also possible to see the additional computational effort for the entropy-conservative flux
differencing DG methods for both physical systems. This is because the numerical fluxes are more
expensive in
terms of computational cost, and require special care to optimize their performance~\cite{hendrik_blog,ranocha2021efficient}.
We reiterate that both codes implement the same numerical nodal DG methods on curvilinear meshes used in these tests.
This demonstrates the suitability of Julia for this kind of simulation-focused scientific computing.

While our Julia code is faster than the mature HPC Fortran code FLUXO for this non-trivial
example, we do not claim that Julia is generally faster than Fortran, C, or C++.
Instead, we would like to emphasize that well-written Julia code can be \emph{at least
as fast} as code written in these traditional scientific computing languages,
as demonstrated also by several microbenchmarks \cite{bezanson2017julia}.
Further, we expect to be able to achieve similar performance with either language
by spending enough time and effort to optimize the respective codes.
\trixi owes its performance optimizations in part to the code introspection and
profiling tools available in Julia. Similar tools used to optimize the performance in other
languages are often not as easy to use as their Julia counterparts.
